
---
title: "Update CellChatDB by adding user-defined ligand-receptor pairs"
author: "Suoqin Jin"
output: html_document
mainfont: Arial
vignette: >
  %\VignetteIndexEntry{Update CellChatDB by adding user-defined ligand-receptor pairs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  root.dir = './'
)
knitr::opts_chunk$set(eval = FALSE)
```


This vignette outlines the steps to update CellChatDB by adding user-defined ligand-receptor pairs. To do so, the format of the users’ lists must be compatible with the input files of CellChatDB. Users can update CellChatDB by submitting their lists via a pull request at Github (https://github.com/sqjin/CellChat) or following the four steps as follows. 


## Load the required libraries
```{r message=FALSE,warning=FALSE}
library(CellChat)
options(stringsAsFactors = FALSE)
```

## Step 1: Access the ligand-receptor interaction information in CellChatDB
Extract the database information in CellChatDB and then save them in a local computer, including four files: 'geneInfo.csv', 'interaction_input_CellChat.csv', 'complex_input_CellChat.csv', 'and cofactor_input_CellChat.csv'. Users can do it by running the following codes in Rstudio:

```{r}
CellChatDB <- CellChatDB.mouse # set CellChatDB <- CellChatDB.human if working on the human dataset
interaction_input <- CellChatDB$interaction
complex_input <- CellChatDB$complex
cofactor_input <- CellChatDB$cofactor
geneInfo <- CellChatDB$geneInfo
write.csv(interaction_input, file = "interaction_input_CellChatDB.csv")
write.csv(complex_input, file = "complex_input_CellChatDB.csv")
write.csv(cofactor_input, file = "cofactor_input_CellChatDB.csv")
write.csv(geneInfo, file = "geneInfo_input_CellChatDB.csv")

```

## Step 2: Update the required files by adding users’ curated ligand-receptor pairs
Update the above four .csv files by adding users’ curated ligand-receptor pairs. 

* The main file is 'interaction_input_CellChatDB.csv'. Users can first update the ligands, receptors and co-factors in the corresponding columns in 'interaction_input_CellChatDB.csv'. 

* Users can then update "complex_input_CellChatDB.csv" and "cofactor_input_CellChatDB.csv" if any ligand complex, receptor complex and cofactors are updated. Users need to make sure that user-defined complex/cofactor names are the same in 'interaction_input_CellChatDB.csv' and "complex_input_CellChatDB.csv", 'interaction_input_CellChatDB.csv' and " cofactor_input_CellChatDB.csv".

* "geneInfo_input_CellChatDB.csv" contains all gene information in mouse and it should have a column named ‘Symbol’, which does not need to be changed when updating CellChatDB. 


## Step 3: Update CellChatDB
Update CellChatDB once updating the four .csv files. Users can do it by running the following codes in Rstudio:
```{r}
options(stringsAsFactors = FALSE)
interaction_input <- read.csv(file = 'interaction_input_CellChatDB.csv', row.names = 1)
complex_input <- read.csv(file = 'complex_input_CellChatDB.csv', row.names = 1)
cofactor_input <- read.csv(file = 'cofactor_input_CellChatDB.csv', row.names = 1)
geneInfo <- read.csv(file = ' geneInfo_input_CellChatDB.csv', row.names = 1)
CellChatDB <- list()
CellChatDB$interaction <- interaction_input
CellChatDB$complex <- complex_input
CellChatDB$cofactor <- cofactor_input
CellChatDB$geneInfo <- geneInfo

```

## Step 4: Re-build CellChat package (optional)
Re-build CellChat package by updating the database as follows
```{r}
setwd("/Users/$USERS/Downloads/CellChat-master") # This is the folder of CellChat package downloaded from Github
CellChatDB.mouse <- CellChatDB
usethis::use_data(CellChatDB.mouse, overwrite = TRUE)

# If working on a human dataset, do following:
# CellChatDB.human <- CellChatDB
# usethis::use_data(CellChatDB.human, overwrite = TRUE)
```
